Happiness Does Not Cause Altruism: A review of US General Social Survey data

### 2023-03-17 outputs/paper/paper.qmd
Add raw data tables to paper, update attributions

---
title: "Happiness Does Not Cause Altruism: A review of US General Social Survey data"
format: pdf
author: "Sebastian Rodriguez, Randall Ni, Iz Leitch"
abstract: "The relationship between happiness and altruistic behaviors has been the subject of much research, with some studies suggesting a positive correlation and others finding no significant relationship. This paper examines the existing literature on the topic in addition to the survey results provided by NORC’s General Social Survey (GSS) of United States residents and analyzes the data to draw conclusions. Based on the available evidence, it appears that there is little to no relationship between happiness and altruistic behaviors, suggesting that individuals may engage in altruistic acts for reasons other than personal happiness." 
date: today
date-format: long
editor: visual
toc: TRUE
number-sections: TRUE
bibliography: references.bib
thanks: "Code and data supporting this analysis is available at: https://github.com/seb646/happiness-and-altruism"
---

```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Preamble ####
# Title: Happiness Does Not Cause Altruism: A review of US General Social Survey data
# Author: Sebastian Rodriguez, Randall Ni, Iz Leitch
# Email: me@srod.ca
# Date: 16 March 2023
# GitHub: https://github.com/seb646/happiness-and-altruism

@@

### Appendix A: Supplementary Survey
The supplementary survey attached to this project is available through Google Forms: https://forms.gle/Zr2jj93bXFaaPkPdA.

>>>>

---
title: "Happiness Does Not Cause Altruism: A review of US General Social Survey data"
format: pdf
author: "Sebastian Rodriguez, Laura Lee-Chu, Iz Leitch"
abstract: "The relationship between happiness and altruistic behaviors has been the subject of much research, with some studies suggesting a positive correlation and others finding no significant relationship. This paper examines the existing literature on the topic in addition to the survey results provided by NORC’s General Social Survey (GSS) of United States residents and analyzes the data to draw conclusions. Based on the available evidence, it appears that there is little to no relationship between happiness and altruistic behaviors, suggesting that individuals may engage in altruistic acts for reasons other than personal happiness." 
date: today
date-format: long
editor: visual
toc: TRUE
number-sections: TRUE
bibliography: references.bib
thanks: "Code and data supporting this analysis is available at: https://github.com/seb646/happiness-and-altruism"
---

```{r}
#| messages: false
#| echo: false
#| warning: false
#| output: false
#### Preamble ####
# Title: Happiness Does Not Cause Altruism: A review of US General Social Survey data
# Author: Sebastian Rodriguez, Laura Lee-Chu, Iz Leitch
# Email: me@srod.ca
# Date: 16 March 2023
# GitHub: https://github.com/seb646/happiness-and-altruism

@@

### Appendix A: Raw Data

#### Happiness
```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Raw GSS Data: General Happiness (by year)"
#| label: fig-5
#### Create a table summarizing happiness data from GSS ####
summarized_happiness_data <- merge(x = happy_positive_data, y = happy_negative_data, by = "year", all.x = TRUE) |>
   # Rename columns
  rename("Year" = "year", '"Very" or "Pretty" happy' = "people.x", '"Not too happy"' = "people.y")

summarized_happiness_data |>
  # Table configuration
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>

  # Add headers for table data
  add_header_above(
    c(" ", "Number of People Feeling" = 2),
    align = "c",
    bold = TRUE
  ) |>

  # Style the table
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE) |>

  # Make the table headers bold
  row_spec(0, bold = TRUE) |>

  # Increase width of first column
  column_spec(1, width = "10em")
```

#### Donating to the Homeless
```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Raw GSS Data: Giving Food or Money to the Homeless (by year)"
#| label: fig-6
#### Create a table summarizing homeless donation data from GSS ####
summarized_homeless_data <- merge(x = homeless_positive_data, y = homeless_negative_data, by = "year", all.x = TRUE) |>
   # Rename columns
  rename("Year" = "year", 'At least twice a year' = "people.x", 'Less than twice a year' = "people.y")

summarized_homeless_data |>
  # Table configuration
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>

  # Add headers for table data
  add_header_above(
    c(" ", "Number of People Donating to the Homeless" = 2),
    align = "c",
    bold = TRUE
  ) |>

  # Style the table
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE) |>

  # Make the table headers bold
  row_spec(0, bold = TRUE) |>

  # Increase width of first column
  column_spec(1, width = "10em")
```

#### Giving Directions to a Stanger
```{r}
#| messages: false
#| echo: false
#| warning: false
#| fig-cap: "Raw GSS Data: Giving Directions to a Stanger (by year)"
#| label: fig-7
#### Create a table summarizing giving strangers directions data from GSS ####
summarized_directions_data <- merge(x = directions_positive_data, y = directions_negative_data, by = "year", all.x = TRUE) |>
   # Rename columns
  rename("Year" = "year", 'At least twice a year' = "people.x", 'Less than twice a year' = "people.y")

summarized_directions_data |>
  # Table configuration
  knitr::kable(align = "l", booktabs = TRUE, escape = FALSE, linesep = "") |>

  # Add headers for table data
  add_header_above(
    c(" ", "Number of People Giving Strangers Directions" = 2),
    align = "c",
    bold = TRUE
  ) |>

  # Style the table
  kable_styling(font_size = 10, latex_options = "hold_position", full_width = TRUE) |>

  # Make the table headers bold
  row_spec(0, bold = TRUE) |>

  # Increase width of first column
  column_spec(1, width = "10em")
```

### Appendix B: Supplementary Survey
The supplementary survey attached to this project is available through Google Forms: https://forms.gle/Zr2jj93bXFaaPkPdA.
